### 2014-09-05 paper.tex
Beefed up the raw data section, added a make file, and add more to outline.

@@

@@

\section{Introduction}

Emergig powered prothetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an abled bodied person. A modern gait lab is able to
collect a variety of kinmetic, kinetic, and physiological data from humans
during gait. This data can be used to drive the design of the controller. With
a rich enough data set, one may be able to identify control mechanism used
during a human's natural gait and recovery from perturbations. We have
collected one such data set that we believe is richer than previous gait data
sets and may be rich enough for control identification. The data can also be
used for verifciation purposes for controllers that have been designed in other
manners.

@@

\section{Raw Data}

The raw data consists of a set of ASCII tab delimited text files output from
both the Mocap and Record modules in D-Flow in addition a YAML file that
contains all of the necessary meta data for the given trial. These three files
are stored in a hierarchy of directories with one trial per directory. The
directories are named in the following fashion \verb+T001/+ where \verb+T+
stands for ``trial'' and the following three digits are provide a unique trial
identification number.

@@

\begin{description}
  \item[TimeStamp] The system monotonically time when D-Flow recieves a frame
    from Cortex. These are approximately at 100 hz and given in seconds.
  \item[FrameNumber] Monotonically increasing integers that correspond to each
    frame received from Cortex.
  \item[Marker Coordinates] Any column that ends in \verb+.PosX+, \verb+.PosY+,
    or \verb+.PosZ+ are marker coordinates expressed in Cortex's Cartesian
    reference frame. These values are in meters.
  \item[Ground Reaction Loads] The left \verb+FP1+ and right \verb+FP2+ force
    plates output the forces and moments of the BLANK

Markers

We make use of the full body 47 marker set described in \cite{Ton's HBM paper}.

>>>>

% Editing notes
% - No tabs, just spaces!
% - No hanging whitespace
% - Wrap lines to 79 characters
% - Indentation should be two spaces per indent
% - Don't worry about formatting details at first, just get the content in
%   place.

@@

\usepackage{todo}

@@

\section{Introduction}
%
Emergig powered prothetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an abled bodied person. A modern gait lab is able to
collect a variety of kinmetic, kinetic, and physiological data from humans
during gait. This data can be used to drive the design of the controller. With
a rich enough data set, one may be able to identify control mechanism used
during a human's natural gait and recovery from perturbations. We have
collected one such data set that we believe is richer than previous gait data
sets and may be rich enough for control identification. The data can also be
used for verifciation purposes for controllers that have been designed in other
manners.

@@

\section{Equipment}
%
The Human Motion and Control Lab at Cleveland State University has such a
system. Our system includes:

\begin{itemize}
  \item A ForceLink R-Mill which has dual 6 DoF force plates, independent belts
    for each foot, and lateral and pitch motion capabilities.
  \item A 10 Camera Motion Analysis motion capture system which includes the
    Cortex software and hardware for collecting analog and camera data
    simultaneously.  Delsys wireless EMG + 3D Accelerometers.
  \item Motek Medical’s D-Flow software and visual display system.
\end{itemize}

Cortex alone is capable of delivering data from the cameras, force plates, and
analog sensors (EMG/Accelerometer), but D-Flow is required to collect data from
digital sensors and the treadmill’s motion (lateral, pitch, and belts). D-Flow
can output multiple types of files which contain the different data.

Our motion capture system’s coordinate system is such that the X coordinate
points to the right, the Y coordinate points upwards, and the Z coordinate
follows from the right-hand-rule, i.e. points backwards with respect to a
person walking forward on the treadmill. The camera’s coordinate system is
aligned to an origin point on treadmill’s surface during camera calibration.

\section{Raw Data}
%
The raw data consists of a set of ASCII tab delimited text files output from
both the Mocap and Record modules in D-Flow in addition to a YAML file that
contains all of the necessary meta data for the given trial. These three files
are stored in a hierarchy of directories with one trial per directory. The
directories are named in the following fashion \verb+T001/+ where \verb+T+
stands for ``trial'' and the following three digits are provide a unique trial
identification number.

The text file output from the mocap module in DFlow is a tab delimited file.
The first line is the header and contains a time stamp column, frame number
column, marker position columns, force plate force/moment columns, force plate
center of pressure columns, other analog columns, and potentially results from
the real time Human Body Model \cite{Bogert2013} which is included with the
D-Flow software. These are detailed below. The numerical values of the
measurements are provided in decimal floating point notation with 6 decimals of
precision, e.g. 123456.123456 [\%1.6f].

@@

\begin{description}
  \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
    recieves a frame from Cortex. These are approximately at 100 hz and given
    in seconds.
  \item[FrameNumber] Monotonically increasing positive integers that correspond
    to each frame received from Cortex.
  \item[Marker Coordinates] Any column that ends in \verb+.PosX+, \verb+.PosY+,
    or \verb+.PosZ+ are marker coordinates expressed in Cortex's Cartesian
    reference frame. These values are in meters.
  \item[Ground Reaction Loads] There are three ground reaction forces and three
    ground reaction moments recorded by each of the two force plates in Newtons
    and Newton-Meters, respectively. The prefix for these columns is either
    \verb+FP1+ or \verb+FP2+ and represents either force plate 1 (left) or 2
    (right). The suffixes are either \verb+.For[XYZ]+, \verb+.Mom[XYZ]+ for the
    forces and moments, respectively. The force plate voltages are sampled at a
    much higher frequency than the cameras, but delivered at the Cortex camera
    sample rate, 100 Hz through the D-Flow mocap module. A force/moment
    calibration matrix stored in Cortex converts the voltages to forces and
    moments before sending it to D-Flow [1]. Cortex also computes the center of
    pressure from the forces, moments, and force plate dimensions. These have
    the same prefixes for the plate number, have the suffix \verb+.Cop[XYZ]+,
    and are in meters.
  \item[Analog Channels] Several analog signals are recorded under column
    headers \verb+Channel[1-99].Anlg+. These correspond to analog signals
    sampled by Cortex and correspond to the ?92? National Instruments XXXX DAQ
    box. The first twelve are the voltages from the force plate load cells. We
    also record the acceleration of 4 points on the treadmill base in chanels
    X-X.
\end{description}

\subsection{record-xxx.txt}
%
The record module also outputs a tab delimited ASCII text file. It includes a
\verb+Time+ column which records the D-Flow system time in seconds which
corresponds to the time recorded in the \verb+TimeStamp+ column in mocap module
tsv file. There are two additional columns \verb+RightBeltSpeed+ and
\verb+LeftBeltSpeed+ which provide the independent belt speeds measure in
meters per second by an ?encoder?.

Finally, the record module is capable of recording the time at which various
events occur, as detected or set by D-Flow. It does this by inserting commented
(\#) lines in between the rows when the event occurred. The record files have
several events that correspond to the phases of the protocol:

\begin{description}
  \item[A: Force Plate Zeroing] Marks the time at which there is no load on the
    force plates and when the voltages were zeroed.
  \item[B: Calibration Pose] Marks the time at which the person is in the
    T-Pose.
  \item[C: First Normal Walking] Marks the time when the treadmill begins Phase
    1: constant belt speed.
  \item[D: Longitudinal Perturbation] Marks the time when the treadmill begins
    Phase 2: longitudinal perturbations in the belt speed.
  \item[E: Second Normal Walking] Marks the time when phase 3 starts: constant
    belt speed.
  \item[F: Unloaded End] Marks the time at which there is no load on the force
    plates and the belts are stationary.
\end{description}

\subsection{meta-xxx.yml}

Each trial directory contains a meta data file in the YML format. There are
three main headings: \verb+study+, \verb+subject+, and \verb+trial+.

The \verb+study+ section contains identififying informatio for the overall
study, an identification number, name, and description. This is the same for
all meta data files in the study.

The \verb+subject+ provides key value pairs of information about the subject in
that trial. Each subject has a unique identification number and basic
anthropomorphic data is included.

The \verb+trial+ section contains the unique information about the trial. Each
trial has a unique identification number.

\todo[inline]{Add and example meta data file here with explanations of each
field.}

\subsection{Markers}
%
We make use of the full body 47 marker set described in \cite{Ton's HBM paper}
and presented in detail in Table \ref{tab:marker-labels}. As with all camera
based motion capture systems, the markers sometimes go missing in the
recording.

If the data was recorded in a D-Flow version less than 3.16.2rc4 [3], D-Flow
records the last non-missing value in all three axes until the marker is
visible again when a marker goes missing.

In D-Flow versions greater than or equal to 3.16.2rc4 the missing markers are
indicated in the TSV file as either 0.000000 or -0.000000, which is the same as
has been in the HBM columns in all versions of D-Flow. The D-Flow version must
be provided in the meta data yml file, otherwise it will assume D-Flow is at
the latest version.

\begin{table}
  \begin{tabular}{rlll}
    1 & LHEAD & Left head & Just above the ear, in the middle. \\
    2 & THEAD & Top head & On top of the head, in line with the LHEAD and RHEAD. \\
    3 & RHEAD & Right head & Just above the ear, in the middle. \\
    4 & FHEAD & Forehead & Between line LHEAD/RHEAD and THEAD a bit right from center. \\
    5 & C7 & C7 & On the 7th cervical vertebrae. \\
    6 & T10 & T10 & On the 10th thoracic vertbrae. \\
    7 & SACR & Sacrum bone & on the scacral bone. \\
    8 & NAVE & Navel & On the navel. \\
    9 & XYPH & Xiphoid process & Xiphoid process of the sternum. \\
    10 & STRN & Sternum & On the jugular notch of the sternum. \\
    11 & BBAC & Scapula & On the inferior angle fo the right scapular. \\
    12 & LSHO & Left shoulder & Left acromion. \\
    13 & LDELT & Left deltoid muscle & Apex of the deltoid muscle. \\
    14 & LLEE & Left lateral elbow & Left lateral epicondyle of the elbow. Upper one in the T-Pose. \\
    15 & LMEE & Left medial elbow & Left medial epicondyle of the elbow. Lower on in the T-Pose. \\
    16 & LFRM & Left forearm & On 2/3 on the line between the LLEE and LMW. \\
    17 & LMW & Left medial wrist & On styloid process radius, thumb side. \\
    18 & LLW & Left lateral wrist & On styloid process ulna, pinky side. \\
    19 & LFIN & Left fingers & Center of the hand. Caput metatarsal 3. \\
    20 & RSHO & Right shoulder & Right acromion. \\
    21 & RDELT & Right deltoid muscle & Apex of deltoid muscle. \\
    22 & RLEE & Right lateral elbow & Right lateral epicondyle of the elbow. Lower one in the T-pose. \\
    23 & RMEE & Right medial elbow & Right medial epicondyle of the elbow. Lower one in the T-pose. \\
    24 & RFRM & Right forearm & On 1/3 on the line between the RLEE and RMW. \\
    25 & RMW & Right medial wrist & On styloid process radius, thumb side. \\
    26 & RLW & Right lateral wrist & On styloid process ulna, pinky side. \\
    27 & RFIN & Right fingers & Center of the hand. Caput metatarsal 3. \\
    28 & LASIS & Pelvic bone left front & Left anterio superior iliac spine. \\
    29 & RASIS & Pelvic bone right front & Right anterior superior iliac spine. \\
    30 & LPSIS & Pelvic bone left back & Left posterior superio iliac spine. \\
    31 & RPSIS & Pelvic bone right back & Right posterior superior iliac spine. \\
    32 & LGTRO & Left greater trochanter of the femur & On the cetner of the left greater trochanter. \\
    33 & FLTHI & Left thigh & On 1/3 on the line between the LFTRO and LLEK. \\
    34 & LLEK & Left lateral epicondyle of the knee & On the lateral side of the joint axis. \\
    35 & LATI & Left anterior of the tibia & On 2/3 on the line between the LLEK and LLM. \\
    36 & LLM & Left lateral malleoulus of the ankle & The center of the heel at the same height as the toe. \\
    37 & LHEE & Left heel & Center of the heel at the same height as the toe. \\
    38 & LTOE & Left toe & Tip of big toe. \\
    39 & LMT5 & Left 5th metatarsal & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
    40 & RGTRO & Right greater trochanter of the femur & On the cetner of the right greater trochanter. \\
    41 & FRTHI & Right thigh & On 2/3 on the line between the RFTRO and RLEK. \\
    42 & RLEK & Right lateral epicondyle of the knee & On the lateral side of the joint axis. \\
    43 & RATI & Right anterior of the tibia & On 1/3 on the line between the RLEK and RLM. \\
    44 & RLM & Right lateral malleoulus of the ankle & The center of the heel at the same height as the toe. \\
    45 & RHEE & Right heel & Center of the heel at the same height as the toe. \\
    46 & RTOE & Right toe & Tip of big toe. \\
    47 & RMT5 & Right 5th metatarsal & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
  \end{tabular}
  \label{tab:marker-labels}
\end{table}
